\documentclass[11pt]{article}
\usepackage[width=7.5in, height=9.5in, top=0.5in, papersize={8.5in,11in}]{geometry}
\usepackage{ulem}
\usepackage[dvipsnames,svgnames]{xcolor}
\usepackage[pdftex]{graphicx}
\title{SDSMT SENIOR DESIGN SOFTWARE DEVELOPMENT AGREEMENT}
\usepackage{hyperref}
\setlength{\parskip}{2mm}
\setlength{\parindent}{0mm}


\begin{document}
\typeout{"This is the Software Agreement.  Take this seriously!"}


{\Large \bf 
\centerline{SDSMT SENIOR DESIGN}\centerline{SOFTWARE DEVELOPMENT AGREEMENT}
}
\vspace{\baselineskip}

This Software Development Agreement (the ``Agreement") is made between the SDSMT  Computer Science\\[3mm] Senior Design Team  \hrulefill, \\[-1mm]  \hspace*{9cm} {\footnotesize (``Student Group")}\\[3mm]
consisting of team members  \hrulefill,  \\[-1mm]  \hspace*{9cm} {\footnotesize (``Student Names")}\\[3mm]
 and  Sponsor \hrulefill,  \\[-1mm]  \hspace*{9cm}{\footnotesize  (``Company Name")} \\[3mm]
 with address: \hrulefill . 

[Note: Bracketed material is included to suggest content that will vary with each agreement. I STRONGLY SUGGEST THAT THE INSTRUCTOR LOOK AT THE COMPLETED AGREEMENT BEFORE YOU SIGN IT!!  ]  \vspace{4mm}


\section{RECITALS}
\begin{enumerate}  \itemsep4pt \parskip0pt \parsep0pt
\item Sponsor desires Senior Design Team to develop software [for use in       Sponsor's simulation platform for optical fiber transmissions of       digitized video signals] (the "Field").    

\item Senior Design Teams willing to develop such Software.  
\end{enumerate}
NOW, THEREFORE, in consideration of the mutual covenants and promises herein contained, the Team and Sponsor agree as follows:  

\section{EFFECTIVE DATE }

This Agreement shall be effective as of \rule{4cm}{0.4pt}  (the ``Effective Date").  

\section{DEFINITIONS }
\begin{enumerate}  \itemsep4pt \parskip0pt \parsep0pt
\item ``Software" shall mean [the computer programs in machine readable       object code form and any subsequent error corrections or updates       supplied to Sponsor by Senior Design Team pursuant to this Agreement.]        [Depending on the particulars of each agreement, any or all of the       following may need to be specified. If they are relevant, they       should be used throughout, modifying the standard form as       appropriate.]     

\item ``Acceptance Criteria" means the written technical and operational       performance and functional criteria and documentation standards       set out in the [design document.]    

\item ``Acceptance Date" means [the date (prior to the design fair) for each Milestone when all       Deliverables included in that Milestone have been accepted by       Sponsor in accordance with the Acceptance Criteria and this       Agreement.]    

\item  ``Deliverable" means a deliverable specified in the [design document deliverable section.]    

\item  ``Delivery Date" shall mean, [with respect to a particular       Milestone,] the date on which University has delivered to Sponsor       all of the Deliverables [for that Milestone] in accordance with       [the design document and] this Agreement.    [This must be before the end of the semester.]

\item  ``Documentation" means the documents, manuals and written materials       (including end-user manuals) referenced, indicated or described       in [the design document] or otherwise developed pursuant to this       Agreement.    

\item  ``Milestone" means the completion and delivery of all of the       Deliverables or other events which are included or described in       [design document] scheduled for delivery and/or completion on a       given target date; a Milestone will not be considered completed       until the Acceptance Date has occurred with respect to all of the       Deliverables for that Milestone.  
\end{enumerate}

\section{DEVELOPMENT OF SOFTWARE }
\begin{enumerate}  \itemsep4pt \parskip0pt \parsep0pt
\item  Senior Design Team will use its best efforts to develop the Software described in [design document deliverable section.] The Software development will be under the direction of  or his/her successors as mutually agreed to by the parties (``Team Lead") and will be conducted by the Team Lead.  The Team will deliver the Software to the satisfaction of the course instructor that reasonable effort has been made to address the needs of the client.  The Team understands that failure to deliver the Software is grounds for failing the course. 

\item  Sponsor understands that the Senior Design course's mission is education and advancement of knowledge, and, consequently, the development of Software must further that mission. The Senior Design Course does not guarantee specific results or any results, and the Software will be developed only on a best efforts basis.  The Software is considered PROOF OF CONCEPT only and is NOT intended for commercial, medical, mission critical or industrial applications.

\item  The Senior Design instructor will act as mediator between Sponsor and Team; and resolve any conflicts that may arise.
\end{enumerate}

\section{COMPENSATION}

[This is entirely subject to negotiation. Normally NO COMPENSATION occurs in a Senior Design Project.  On occasion an intern status and wage is appropriate. ] 

\section{CONSULTATION AND REPORTS}     
\begin{enumerate}  \itemsep4pt \parskip0pt \parsep0pt
\item  Sponsor's designated representative for consultation and       communications with the Team Lead shall be \\[3mm]  \rule{7cm}{0.4pt}    or such other person as Sponsor       may from time to time designate to the Team Lead (``Designated Representative").    

\item During the Term of the Agreement, Sponsor's representatives may       consult informally with course instructor regarding the       project, both personally and by telephone. Access to work carried       on in University facilities, if any, in the course of this Agreement shall       be entirely under the control of University personnel but shall be       made available on a reasonable basis.    

\item The Team Lead will submit written progress reports. At the conclusion of this Agreement, the Team Lead shall submit a comprehensive final report in the form of the formal course documentation at the conclusion of the Senior Design II course. 
\end{enumerate}

\section{CONFIDENTIAL INFORMATION }    
\begin{enumerate}  \itemsep4pt \parskip0pt \parsep0pt
\item The parties may wish, from time to time, in connection with work       contemplated under this Agreement, to disclose confidential       information to each other (``Confidential Information"). Each party       will use reasonable efforts to prevent the disclosure of any of       the other party's Confidential Information to third parties for a       period of three (3) years after the termination of this Agreement,       provided that the recipient party's obligation shall not apply to       information that:               
\begin{enumerate}  \itemsep4pt \parskip0pt \parsep0pt
\item  is not disclosed in writing or reduced to writing and so                 marked with an appropriate confidentiality legend within                 thirty (30) days of disclosure;              
\item is already in the recipient party's possession at the                 time of disclosure thereof;              
\item  is or later becomes part of the public domain through no                 fault of the recipient party;              
\item  is received from a third party having no obligations of                 confidentiality to the disclosing party;              
\item is independently developed by the recipient party; or              
\item is required by law or regulation to be disclosed.     
\end{enumerate}
\item In the event that information is required to be disclosed pursuant       to subsection (6), the party required to make disclosure shall       notify the other to allow that party to assert whatever exclusions       or exemptions may be available to it under such law or regulation.  
\end{enumerate}


\section{INTELLECTUAL PROPERTY RIGHTS } 

[Negotiated on a case-by-case basis. This must address who owns the algorithms and who owns the source code. For example: All deliverables become property of the Sponsor. Roughly:  If the idea originates with the sponsor, or if a sponsor pays you to develop an idea, then they have legitimate claim to the IP.  If the idea originates from the University (through faculty or staff) then the University has legitimate claim. If the idea is yours (student) and you develop it without external compensation then you have legitimate claim. ] 

\section{WARRANTIES }  

The Senior Design Team represents and warrants to Sponsor that:         
\begin{enumerate}  \itemsep4pt \parskip0pt \parsep0pt
\item  the Software is the original work of the Senior Design Team in each and all aspects;        

\item the Software and its use do not infringe any copyright or   trade secret rights of any third party.  
\end{enumerate}
No agreements will be made beyond items (1) and (2).

\section{INDEMNITY}   
\begin{enumerate}  \itemsep4pt \parskip0pt \parsep0pt
\item Sponsor is responsible for claims and damages, losses or expenses held against the Sponsor. [Sponsor may have something to add here.]

\item  Sponsor shall       indemnify and hold harmless the Senior Design Team, its affiliated companies       and the officers, agents, directors and employees of the same from       any and all claims and damages, losses or expenses, including       attorney's fees, caused by any negligent act of Sponsor or any of       Sponsor's agents, employees, subcontractors, or suppliers.    

\item  NEITHER PARTY TO THIS AGREEMENT NOR THEIR AFFILIATED COMPANIES,       NOR THE OFFICERS, AGENTS, STUDENTS AND EMPLOYEES OF ANY OF THE       FOREGOING, SHALL BE LIABLE TO ANY OTHER PARTY HERETO IN ANY ACTION       OR CLAIM FOR CONSEQUENTIAL OR SPECIAL DAMAGES, LOSS OF PROFITS,       LOSS OF OPPORTUNITY, LOSS OF PRODUCT OR LOSS OF USE, WHETHER THE       ACTION IN WHICH RECOVERY OF DAMAGES IS SOUGHT IS BASED ON CONTRACT       TORT (INCLUDING SOLE, CONCURRENT OR OTHER NEGLIGENCE AND STRICT       LIABILITY), STATUTE OR OTHERWISE. TO THE EXTENT PERMITTED BY LAW,       ANY STATUTORY REMEDIES WHICH ARE INCONSISTENT WITH THE PROVISIONS       OF THESE TERMS ARE WAIVED.  
\end{enumerate}

\section{INDEPENDENT CONTRACTOR}  

For the purposes of this Agreement and all services to be provided hereunder, the parties shall be, and shall be deemed to be, independent contractors and not agents or employees of the other party. Neither party shall have authority to make any statements, representations or commitments of any kind, or to take any action which shall be binding on the other party, except as may be expressly provided for herein or authorized in writing.  

\section{TERM AND TERMINATION }    
\begin{enumerate}  \itemsep4pt \parskip0pt \parsep0pt
\item This Agreement shall commence on the Effective Date and extend until the end of classes of the second semester of Senior Design (CSC 467), unless sooner terminated in accordance with the provisions of this Section (``Term"). 

\item This Agreement may be terminated by the written agreement of both       parties.    

\item In the event that either party shall be in default of its       materials obligations under this Agreement and shall fail to       remedy such default within thirty (30) days after receipt of       written notice thereof, this Agreement shall terminate upon       expiration of the thirty (30) day period.    

\item Any provisions of this Agreement which by their nature extend       beyond termination shall survive such termination.  
\end{enumerate}


\section{ATTACHMENTS}  

Attachments A and B are incorporated and made a part of this Agreement for all purposes.  

\section{GENERAL }    
\begin{enumerate}  \itemsep4pt \parskip0pt \parsep0pt
\item This Agreement constitutes the entire and only agreement between       the parties relating to the Senior Design Course, and all prior       negotiations, representations, agreements and understandings are       superseded hereby. No agreements altering or supplementing the       terms hereof may be made except by means of a written document       signed by the duly authorized representatives of the parties.    

\item This Agreement shall be governed by, construed, and enforced in       accordance with the internal laws of the State of South Dakota. 
\end{enumerate}

\section{SIGNATURES}    
\begin{tabular}{ll}
  \strut\vspace{0.25in} & \\
  \makebox[3in]{\hrulefill} & \makebox[2in]{\hrulefill} \\
  Replace with name of student \#1 & Date \\
  \strut\vspace{0.25in} & \\
  \makebox[3in]{\hrulefill} & \makebox[2in]{\hrulefill} \\
  Replace with name of student \#2 & Date \\
  \strut\vspace{0.25in} & \\
  \makebox[3in]{\hrulefill} & \makebox[2in]{\hrulefill} \\
  Replace with name of student \#3 & Date \\
  \strut\vspace{0.25in} & \\
  \makebox[3in]{\hrulefill} & \makebox[2in]{\hrulefill} \\
  Replace with name of sponsor's representative & Date \\
\end{tabular}

\end{document}
